---
title: "Device_Comparison"
date: "`r Sys.Date()`"
output: html_document
---
This is the analysis file of the device comparison (Empatica vs Polar) of physiological data (i.e. HR & HRV) of the study \_\_\_\_\_\_\_. For further information on the study, please refer to the corresponding paper.
---

# 0. Packages

a)  Download Data Wrangling Packages

```{r data-wrangling-packages,include=FALSE	}

library(readxl) #To import the data
library(tidyverse) #To prepare the data adequately
library(magrittr) #To ensure seamless workflow
library(writexl) #To export data
library(purrr) #programming tools 
library(stringr) #wrappers

```

b)  Download Statistics & Analyis Packages

```{r statistics-packages, include=FALSE	}

library(rstatix)#Pipe-Friendly Framework for Basic Statistical Tests 
library(psych) #A package for personality, psychometric, and psychological research
library(effsize) #Efficient Effect Size Computation
library (Hmisc) #Harrell Miscellaneous
library(jtools) #analysis and presentation of data
library(knitr) #reporting 
library(ggcorrplot) #visualization of a correlation matrix
library(patchwork)
library(afex) #analysis of factorial experiments 
library(lme4) #mixed models 
library(lmerTest) #mixed models
library(emmeans) #estimated marginal means
library(performance) #check regression models
library(easystats)

```

c)  TimeSeries and HR Data Packages

```{r timeseries-packages, include=FALSE	}

library(imputeTS) #timeseries imputation
library(RHRV) #HRV analysis tool
library(slider) #sliding window function
library(pracma) #for imputations (math function package)
library(hms) #for pretty time of day
library(lubridate) #parse & manipulate dates 
library(BlandAltmanLeh) #Bland Altman Plot for Agreement Comparison

```

# 1. Import 

## 1.1 Import: Physiological Data

```{r import-hrv-data}

#Set WD
setwd("./Preprocessed_Data")

#Preprocessed Dataset Cardio_Comparison 
Cardio_Comparison <- read_excel("Cardio_Comparison.xlsx")
Cardio_Comparison$Event_Phase <- factor(
  Cardio_Comparison$Event_Phase,
  levels = c("Baseline", "Trials", "Drone"))


#Imputed TimeSeries IBI Data 

## Empatica
Empatica_RR_Imputed <- read_excel("Empatica_RR_Imputed.xlsx")
Empatica_RR_Imputed$Event_Phase <- factor(
  Empatica_RR_Imputed$Event_Phase,
  levels = c("Baseline", "Trials", "Drone"))

## Polar
Polar_RR_Imputed <- read_excel("Polar_RR_Imputed.xlsx")
Polar_RR_Imputed$Timestamp <- as_hms(Polar_RR_Imputed$Timestamp)
Polar_RR_Imputed$Event_Phase <- factor(
  Polar_RR_Imputed$Event_Phase,
  levels = c("Baseline", "Trials", "Drone"))

# Imputed TimeSeries HR Data 

## Empatica
Empatica_HR <- read_excel("Empatica_HR.xlsx")
Empatica_HR$Event_Phase <- factor(
  Empatica_HR$Event_Phase,
  levels = c("Baseline", "Trials", "Drone"))

## Polar
Polar_HR <- read_excel("Polar_HR.xlsx")
Polar_HR$Timestamp <- as_hms(Polar_HR$Timestamp)
Polar_HR$Event_Phase <- factor(
  Polar_HR$Event_Phase,
  levels = c("Baseline", "Trials", "Drone"))

## Computed 1-minute HFn Data 
Empatica_HFn_1min <- read_excel("HFn_1min_empatica.xlsx")
Polar_HFn_1min <- read_excel("HFn_1min_polar.xlsx")

```

#1.2 Import Socio-Data

```{r import-socio-data}

#Set WD
setwd("../Supplementary_Inputs")

#Import Complete Datset
socio <- read_excel("socio.xlsx")


## Recode Factors 
socio$Gender <- dplyr::recode(
  socio$Gender,
  `1` = "male",
  `2` = "female")

```


# 2. Sample Analysis

## 2.1 Sample Description 

```{r sample-description}

#Group Distribution
socio %>% group_by(Group) %>% count()

#Age 
socio %>% 
  summarise(Age_Mean = mean(Age, na.rm=T), 
            Age_SD = sd (Age, na.rm=T),
            min = min(Age, na.rm = T), 
            max = max(Age, na.rm = T))

#Gender 
socio %>% group_by(Gender) %>% count()

```

## 2.1 Sample Size 

```{r sample-size}

#Full Sample Size 
n_distinct(Cardio_Comparison$ID) #45

## Sample Size for each device
Cardio_Comparison %>%
  summarise(
    N_Empatica = n_distinct(ID[!is.na(RMSSD_Empatica)]),
    N_Polar    = n_distinct(ID[!is.na(RMSSD_Polar)])
  )

## Sample Size for IBI-based Analysis
Cardio_Comparison %>%
  group_by(ID) %>%
  summarise(
    Empatica = any(!is.na(N_RR_Empatica)),
    Polar    = any(!is.na(N_RR_Polar)),
    .groups = "drop"
  ) %>%
  mutate(
    Data = case_when(
      Empatica & Polar  ~ "Both",
      Empatica & !Polar ~ "Empatica only",
      !Empatica & Polar ~ "Polar only",
      TRUE              ~ "Neither"
    )
  )  %>% count(Data)

## Sample Size for HR-based Analysis
Cardio_Comparison %>%
  group_by(ID) %>%
  summarise(
    Empatica = any(!is.na(HR_Empatica)),
    Polar    = any(!is.na(HR_Polar)),
    .groups = "drop"
  ) %>%
  mutate(
    Data = case_when(
      Empatica & Polar  ~ "Both",
      Empatica & !Polar ~ "Empatica only",
      !Empatica & Polar ~ "Polar only",
      TRUE              ~ "Neither"
    )
  )  %>% count(Data)

```

# 3. Signal Level Analysis

## 3.1 Artifact Rate: Empatica 

### Overall Artifact Rate 

```{r artifact-rate-empatica}

#Total Artifacts 
Empatica_RR_Imputed %>% 
  count(Artifact_Flag)


#Artifact percentage per participant and event
artifact_empatica <- Empatica_RR_Imputed %>%
  group_by(ID, Event_Phase) %>%
  summarise(
    n_total = n(),
    n_artifact = sum(Artifact_Flag == "Artifact", na.rm = TRUE),
    artifact_percent = (n_artifact / n_total) * 100,
    .groups = "drop"
  )

#Mean and SD across participants for each event
artifact_empatica %>%
  group_by(Event_Phase) %>%
  summarise(
    N = n(),
    Mean_Artifact_Percent = round(mean(artifact_percent, na.rm = TRUE),2),
    SD_Artifact_Percent = round(sd(artifact_percent, na.rm = TRUE),2),
    .groups = "drop"
  ) |> view()

```

### Artifact Ratio per Recording Chunk

- Standard Time-based Measures: RMSSD, SDNN -> up to 36%

- Frequency-based Measures: LF, HF -> up to 2%

```{r artifact-ratio-empatica}

#Define thresholds
thresholds <- c(2, 36)

#Compute recording chunks above artifact ratio threshold
Critical_Summary_Empatica <- artifact_empatica %>%
  crossing(Threshold = thresholds) %>%
  mutate(
    Critical_Value = case_when(
      artifact_percent > Threshold ~ "Critical",
      TRUE ~ "OK")) %>%
  group_by(Threshold, Event_Phase, Critical_Value) %>%
  summarise( n = n(),
             .groups = "drop") %>%
  group_by(Threshold , Event_Phase) %>%
  mutate(
    percentage = round(100 * n / sum(n), 2)) %>%
  ungroup() |> view()

## Visualize
ggplot(Critical_Summary_Empatica,
       aes(x = Event_Phase,
           y = n,
           fill = Critical_Value)) +
  
  geom_col(position = "dodge") +
  
  facet_wrap(
    ~ Threshold,
    labeller = as_labeller(
      c(
        "2" = "Threshold: 2%",
        "36" = "Threshold: 36%")))  +
  
  labs(
    title = "Empatica: Artifact Ratios Across Thresholds",
    x = "Event Phase",
    y = "Number of Observations",
    fill = "Artifact Status"
  )  + 
  theme_minimal()

```

## 3.2 Artifact Rate: Polar RR 

### Overall Artifact Rate 

```{r artifact-rate-polar-rr}

#Total Artifacts 
Polar_RR_Imputed %>% 
  count(Artifact_Flag)

#Artifact percentage per participant and event
artifact_polar_rr <- Polar_RR_Imputed %>%
  group_by(ID, Event_Phase) %>%
  summarise(
    n_total = n(),
    n_artifact = sum(Artifact_Flag == "Artifact", na.rm = TRUE),
    artifact_percent = (n_artifact / n_total) * 100,
    .groups = "drop"
  )

#Mean and SD across participants for each event
artifact_polar_rr %>%
  group_by(Event_Phase) %>%
  summarise(
    N = n(),
    Mean_Artifact_Percent = round(mean(artifact_percent, na.rm = TRUE),2),
    SD_Artifact_Percent = round(sd(artifact_percent, na.rm = TRUE),2),
    .groups = "drop"
  ) |> view()


```

### Artifact Ratio per Recording Chunk

```{r artifact-ratio-polar-rr}

#Compute recording chunks above artifact ratio threshold
Critical_Summary_Polar <- artifact_polar_rr %>%
  crossing(Threshold = thresholds) %>%
  mutate(
    Critical_Value = case_when(
      artifact_percent > Threshold ~ "Critical",
      TRUE ~ "OK")) %>%
  group_by(Threshold, Event_Phase, Critical_Value) %>%
  summarise( n = n(),
             .groups = "drop") %>%
  group_by(Threshold , Event_Phase) %>%
  mutate(
    percentage = round(100 * n / sum(n), 2)) %>%
  ungroup() |> view()

## Visualize 
ggplot(Critical_Summary_Polar,
       aes(x = Event_Phase,
           y = n,
           fill = Critical_Value)) +
  
  geom_col(position = "dodge") +
  
  facet_wrap(
    ~ Threshold,
    labeller = as_labeller(
      c(
        "2" = "Threshold: 2%",
        "36" = "Threshold: 36%")))  +
  
  labs(
    title = "Polar: Artifact Ratios Across Thresholds",
    x = "Event Phase",
    y = "Number of Observations",
    fill = "Artifact Status"
  )  + 
  theme_minimal()

```

## 3.4 N-IBI / N-RR Comparison

```{r ibi-intervall-detection-comparison}

#Plot Count Comparison
Cardio_Comparison %>%
  ggplot(aes(
    x = N_RR_Polar,
    y = N_RR_Empatica,
    color = Event_Phase
  )) +
  geom_point(size = 3, alpha = 0.8) +
  geom_abline(
    slope = 1,
    intercept = 0,
    linetype = "dashed"
  ) +
  facet_wrap(~ ID) +
  labs(
    title = "IBI Count Comparison",
    x = "Polar RR Count",
    y = "Empatica IBI Count"
  ) +
  theme_apa()

#Compute RR_Ratio for all Events 
RR_ratio <- Cardio_Comparison %>%
  mutate(RR_ratio = N_RR_Empatica /N_RR_Polar) %>% group_by(Event_Phase) %>%  
  summarise(RR_ratio = round(mean(RR_ratio, na.rm = T),2)) |> view()


#Visualize RR_Ratio per Participant
Cardio_Comparison %>%
  mutate(
    RR_ratio = N_RR_Empatica / N_RR_Polar
  ) %>%
  ggplot(aes(
    x = Event_Phase,
    y = RR_ratio,
    color = Event_Phase
  )) +
  geom_point() +
  facet_wrap(~ ID) +
  theme_apa()


```

## 3.5 Remove ID03 from Empatica Data due to poor data quality 

```{r set-id03-dead}

#Set Empatica values to NA in Preprocessed Dataset
empatica_cols_ID03 <- grep("_Empatica", names(Cardio_Comparison), value = TRUE)
Cardio_Comparison[Cardio_Comparison$ID == "OITC03", 
                    empatica_cols_ID03] <- NA

#Remove ID03 from Imputed Timeseries Dataset
Empatica_RR_Imputed <- Empatica_RR_Imputed %>% filter(ID != "OITC03")
Empatica_HR <- Empatica_HR %>% filter(ID != "OITC03")
Empatica_HFn_1min <- Empatica_HFn_1min %>% filter(ID != "OITC03")

```

# 4. Index-Level Anaylsis

## 4.1 Split-Half-Reliability: For Empatica

### 4.1.1 Prepare Epochs

#### For Empatica RR

a) Create a new grouping variable: Event_Grouped to group Trial1-Trial3 data 
b) Separate each Event into 1 minute-epochs
c) Compute Metrics on 1 minute epochs

```{r prepare-epochs-rr}

#Prepare Dataset

Empatica_RR_Imputed <- Empatica_RR_Imputed %>%
 
  ## Create sequential 1-minute epoch labels within each event
  group_by(
    ID,
    Event_Phase
  ) %>%

  mutate(

    # seconds elapsed since start of this event
    elapsed_sec =
      as.numeric(timestamp) -
      first(as.numeric(timestamp)),

    # epoch label (1-minute bins)
    Epoch = floor(elapsed_sec / 60) + 1

  ) %>%

  ungroup()

#Compute Metrics
Empatica_RR_Epoch_Summary <- Empatica_RR_Imputed %>% 
  group_by(ID, Event_Phase, Epoch) %>%
  summarise(
    
    # Keep participant info
    Session = first(Session),
    Color = first(Color),
    Station = first(Station),
    Group = first(Group), 
    Type = first(Type), 
    
    # SDNN
    SDNN = round(sd(EWMA_RR, na.rm = TRUE), 2),
    
    # log SDNN
    lnSDNN = round(log(
      sd(EWMA_RR, na.rm = TRUE)
    ), 3),
    
    # RMSSD
    RMSSD = round(
      sqrt(
        mean(
          diff(EWMA_RR)^2,
          na.rm = TRUE
        )
      ),
      2
    ),
    
    # log RMSSD
    lnRMSSD = round(log(
      sqrt(
        mean(
          diff(EWMA_RR)^2,
          na.rm = TRUE
        )
      )
    ), 3),
    
    N_RR = n(),
    
    .groups = "drop")  %>% 
  
  filter (N_RR > 10)

```

#### For Empatica HFn

```{r prepare-epochs-hfn-empatica}

#Prepare Dataset

## Create new grouping variable 
Empatica_HFn_1min <- Empatica_HFn_1min %>%
  mutate(
    Event_Phase = case_when(
      Event_Epoch %in% c("Trial1", "Trial2", "Trial3", "Video") ~ "Trials",
      TRUE ~ Event_Epoch) ) 

```


#### For Empatica HR

```{r prepare-epochs-hr}

#Prepare Dataset

Empatica_HR_Epoch <- Empatica_HR %>%
  
  ## Create sequential 1-minute epoch labels within each event
  group_by(
    ID,
    Event_Phase) %>%

  mutate(

    # seconds elapsed since start of this event
    elapsed_sec =
      as.numeric(timestamp) -
      first(as.numeric(timestamp)),

    # epoch label (1-minute bins)
    Epoch = floor(elapsed_sec / 60) + 1

  ) %>%

  ungroup()

#Compute Metrics
Empatica_HR_Epoch_Summary <- Empatica_HR_Epoch %>% 
  group_by(ID, Event_Phase, Epoch) %>%
  summarise(
    
    # Keep participant info
    Session = first(Session),
    Color = first(Color),
    Station = first(Station),
    Group = first(Group), 
    Type = first(Type), 
    
    # HR
    Mean_HR= round(mean(HR, na.rm = TRUE)),
    N_HR = n(),

    .groups = "drop") %>% 
  
  filter (N_HR > 10)

```

### 4.1.2 Compute ICC for Split-Half Reliability 

a) Label odd & even epoch in each event 
b) Compute average logRMSSD for even & odd epochs per event
c) Comoute agreement between even & odd epochs 

#### For Empatica RMSSD

```{r compute-epochs-icc-empatica-rmssd}

#Compute ICC 
ICC_SplitHalf_Empatica_RMSSD <- data.frame()
events <- unique(Empatica_RR_Epoch_Summary$Event_Phase)

for(event in events) {

  # Subset event
  df_event <- subset(
    Empatica_RR_Epoch_Summary,
    Event_Phase == event
  )

  # Create odd/even split
  df_event <- df_event %>%

    mutate(
      Half = ifelse(
        Epoch %% 2 == 0,
        "Even",
        "Odd"
      )
    ) %>%

    # Mean lnRMSSD per participant × half
    group_by(ID, Half) %>%

    summarise(
      lnRMSSD = mean(
        lnRMSSD,
        na.rm = TRUE
      ),
      .groups = "drop"
    ) %>%

    # Wide format: Odd vs Even columns
    pivot_wider(
      names_from = Half,
      values_from = lnRMSSD
    )

  # Keep complete cases
  df_event <- df_event[
    complete.cases(
      df_event$Odd,
      df_event$Even
    ),
  ]

  # Compute ICC if enough participants
  if(nrow(df_event) > 2) {

    icc <- psych::ICC(
      df_event[, c("Odd", "Even")]
    )

    # Extract ICC(A,1)
    icc_row <- icc$results[
  "Single_raters_absolute",
  ]
    icc_value <- round(icc_row["ICC"],2)
    lower_bound <- round(icc_row["lower bound"],2)
    upper_bound <- round(icc_row["upper bound"],2)

    # Save result
    ICC_SplitHalf_Empatica_RMSSD <- rbind(
      ICC_SplitHalf_Empatica_RMSSD,
      data.frame(
        Event = event,
        ICC_A1 = icc_value, 
        Lower = lower_bound, 
        Upper = upper_bound
      )
    )
  }
}

view(ICC_SplitHalf_Empatica_RMSSD)

```

#### For Empatica HFn

```{r compute-epochs-icc-empatica-hfn}

#Compute ICC 
ICC_SplitHalf_Empatica_HFn <- data.frame()

events <- unique(Empatica_HFn_1min$Event_Phase)

for(event in events) {

  # Subset event
  df_event <- subset(
    Empatica_HFn_1min,
    Event_Phase == event
  )

  # Create odd/even split based on 1-minute intervals
  df_event <- df_event %>%

    mutate(
      Half = ifelse(
        Minute %% 2 == 0,
        "Even",
        "Odd"
      )
    ) %>%

    # Mean HFn per participant × half
    group_by(ID, Half) %>%

    summarise(
      HRV_HFn = mean(
        HRV_HFn,
        na.rm = TRUE
      ),
      .groups = "drop"
    ) %>%

    # Wide format: Odd vs Even columns
    pivot_wider(
      names_from = Half,
      values_from = HRV_HFn
    )

  # Keep participants with both odd and even halves
  df_event <- df_event[
    complete.cases(
      df_event$Odd,
      df_event$Even
    ),
  ]

  # Compute ICC if enough participants
  if(nrow(df_event) > 2) {

    icc <- psych::ICC(
      df_event[, c("Odd", "Even")]
    )

    # Extract ICC(A,1): single-rater absolute agreement
    icc_row <- icc$results[
      "Single_raters_absolute",
    ]

    icc_value <- round(
      icc_row["ICC"],
      2
    )

    lower_bound <- round(
      icc_row["lower bound"],
      2
    )

    upper_bound <- round(
      icc_row["upper bound"],
      2
    )

    # Save result
    ICC_SplitHalf_Empatica_HFn <- rbind(
      ICC_SplitHalf_Empatica_HFn,
      data.frame(
        Event = event,
        ICC_A1 = icc_value,
        Lower = lower_bound,
        Upper = upper_bound
      )
    )
  }
}

View(ICC_SplitHalf_Empatica_HFn)

```

#### For Empatica HR

```{r compute-epochs-icc-empatica-hr}

#Compute ICC 
ICC_SplitHalf_Empatica_HR <- data.frame()
events <- unique(Empatica_HR_Epoch_Summary$Event_Phase)

for(event in events) {

  # Subset event
  df_event <- subset(
    Empatica_HR_Epoch_Summary,
    Event_Phase == event
  )

  # Create odd/even split
  df_event <- df_event %>%

    mutate(
      Half = ifelse(
        Epoch %% 2 == 0,
        "Even",
        "Odd"
      )
    ) %>%

    # Mean lnRMSSD per participant × half
    group_by(ID, Half) %>%

    summarise(
      HR = mean(
        Mean_HR,
        na.rm = TRUE
      ),
      .groups = "drop"
    ) %>%

    # Wide format: Odd vs Even columns
    pivot_wider(
      names_from = Half,
      values_from = HR
    )

  # Keep complete cases
  df_event <- df_event[
    complete.cases(
      df_event$Odd,
      df_event$Even
    ),
  ]

  # Compute ICC if enough participants
  if(nrow(df_event) > 2) {

    icc <- psych::ICC(
      df_event[, c("Odd", "Even")]
    )

    # Extract ICC(A,1)
    icc_row <- icc$results[
  "Single_raters_absolute",
  ]
    icc_value <- round(icc_row["ICC"],2)
    lower_bound <- round(icc_row["lower bound"],2)
    upper_bound <- round(icc_row["upper bound"],2)

    # Save result
    ICC_SplitHalf_Empatica_HR <- rbind(
      ICC_SplitHalf_Empatica_HR,
      data.frame(
        Event = event,
        ICC_A1 = icc_value, 
        Lower = lower_bound, 
        Upper = upper_bound
      )
    )
  }
}

view(ICC_SplitHalf_Empatica_HR)
```

## 4.2. Split-Half-Reliability: For Polar

### 4.2.1 Prepare Epochs

#### For Polar RMSSD

```{r prepare-epochs-polar-rmssd}

#Prepare Dataset

Polar_RR_Imputed <- Polar_RR_Imputed %>%
  
  ## Create sequential 1-minute epoch labels within each event
  group_by(
    ID,
    Event_Phase
  ) %>%

  mutate(

    # seconds elapsed since start of this event
    elapsed_sec =
      as.numeric(Timestamp) -
      first(as.numeric(Timestamp)),

    # epoch label (1-minute bins)
    Epoch = floor(elapsed_sec / 60) + 1

  ) %>%

  ungroup()

#Compute Metrics
Polar_RR_Epoch_Summary <- Polar_RR_Imputed %>% 
  group_by(ID, Event_Phase, Epoch) %>%
  
  summarise(
    
    # Keep participant info
    Session = first(Session),
    Color = first(Color),
    Station = first(Station),
    Group = first(Group), 
    Type = first(Type), 
    
    # SDNN
    SDNN = round(sd(EWMA_RR, na.rm = TRUE), 2),
    
    # log SDNN
    lnSDNN = round(log(
      sd(EWMA_RR, na.rm = TRUE)
    ), 3),
    
    # RMSSD
    RMSSD = round(
      sqrt(
        mean(
          diff(EWMA_RR)^2,
          na.rm = TRUE
        )
      ),
      2
    ),
    
    # log RMSSD
    lnRMSSD = round(log(
      sqrt(
        mean(
          diff(EWMA_RR)^2,
          na.rm = TRUE
        )
      )
    ), 3),
    
    N_RR = n(),
    
    .groups = "drop"
  ) %>% 
  filter (N_RR > 10)
```

#### For Polar HFn

```{r prepare-epochs-polar-hfn}

#Prepare Dataset

## Create new grouping variable 
Polar_HFn_1min <- Polar_HFn_1min %>%
  mutate(
    Event_Phase = case_when(
      Event_Epoch %in% c("Trial1", "Trial2", "Trial3", "Video") ~ "Trials",
      TRUE ~ Event_Epoch)) 

```

#### For Polar HR

```{r prepare-epochs-polar-hr}

#Prepare Dataset 

Polar_HR_Epoch <- Polar_HR %>%
  ## Create sequential 1-minute epoch labels within each event
  group_by(
    ID,
    Event_Phase
  ) %>%

  mutate(

    # seconds elapsed since start of this event
    elapsed_sec =
      as.numeric(Timestamp) -
      first(as.numeric(Timestamp)),

    # epoch label (1-minute bins)
    Epoch = floor(elapsed_sec / 60) + 1

  ) %>%

  ungroup()

#Compute Metrics
Polar_HR_Epoch_Summary <- Polar_HR_Epoch %>% 
  group_by(ID, Event_Phase, Epoch) %>%
  
  summarise(
    
    # Keep participant info
    Session = first(Session),
    Color = first(Color),
    Station = first(Station),
    Group = first(Group), 
    Type = first(Type), 
    
    # HR
    Mean_HR= round(mean(HR, na.rm = TRUE)),
    N_HR = n(),

    .groups = "drop"
  )   %>%  
  filter (N_HR > 10)
```

### 4.2.2 Compute ICC for Split-Half Reliability 

#### For Polar RMSSD

```{r compute-epochs-icc-polar-rmssd}

#Compute ICC 
ICC_SplitHalf_Polar_RMSSD <- data.frame()
events <- unique(Polar_RR_Epoch_Summary$Event_Phase)

for(event in events) {

  # Subset event
  df_event <- subset(
    Polar_RR_Epoch_Summary,
    Event_Phase == event
  )

  # Create odd/even split
  df_event <- df_event %>%

    mutate(
      Half = ifelse(
        Epoch %% 2 == 0,
        "Even",
        "Odd"
      )
    ) %>%

    # Mean lnRMSSD per participant × half
    group_by(ID, Half) %>%

    summarise(
      lnRMSSD = mean(
        lnRMSSD,
        na.rm = TRUE
      ),
      .groups = "drop"
    ) %>%

    # Wide format: Odd vs Even columns
    pivot_wider(
      names_from = Half,
      values_from = lnRMSSD
    )

  # Keep complete cases
  df_event <- df_event[
    complete.cases(
      df_event$Odd,
      df_event$Even
    ),
  ]

  # Compute ICC if enough participants
  if(nrow(df_event) > 2) {

    icc <- ICC(
      df_event[, c("Odd", "Even")]
    )

    # Extract ICC(A,1)
    icc_row <- icc$results[
  "Single_raters_absolute",
  ]
  icc_value <- round(icc_row["ICC"],2)
    lower_bound <- round(icc_row["lower bound"],2)
    upper_bound <- round(icc_row["upper bound"],2)

    # Save result
    ICC_SplitHalf_Polar_RMSSD <- rbind(
      ICC_SplitHalf_Polar_RMSSD,
      data.frame(
        Event = event,
        ICC_A1 = icc_value, 
        Lower = lower_bound, 
        Upper = upper_bound 
      )
    )
  }
}

view(ICC_SplitHalf_Polar_RMSSD)
```

#### For Polar HFn

```{r compute-epochs-icc-polar-hfn}

#Compute ICC 
ICC_SplitHalf_Polar_HFn <- data.frame()

events <- unique(Polar_HFn_1min$Event_Phase)

for(event in events) {

  # Subset event
  df_event <- subset(
    Polar_HFn_1min,
    Event_Phase == event
  )

  # Create odd/even split based on 1-minute intervals
  df_event <- df_event %>%

    mutate(
      Half = ifelse(
        Minute %% 2 == 0,
        "Even",
        "Odd"
      )
    ) %>%

    # Mean HFn per participant × half
    group_by(ID, Half) %>%

    summarise(
      HRV_HFn = mean(
        HRV_HFn,
        na.rm = TRUE
      ),
      .groups = "drop"
    ) %>%

    # Wide format: Odd vs Even columns
    pivot_wider(
      names_from = Half,
      values_from = HRV_HFn
    )

  # Keep participants with both odd and even halves
  df_event <- df_event[
    complete.cases(
      df_event$Odd,
      df_event$Even
    ),
  ]

  # Compute ICC if enough participants
  if(nrow(df_event) > 2) {

    icc <- psych::ICC(
      df_event[, c("Odd", "Even")]
    )

    # Extract ICC(A,1): single-rater absolute agreement
    icc_row <- icc$results[
      "Single_raters_absolute",
    ]

    icc_value <- round(
      icc_row["ICC"],
      2
    )

    lower_bound <- round(
      icc_row["lower bound"],
      2
    )

    upper_bound <- round(
      icc_row["upper bound"],
      2
    )

    # Save result
    ICC_SplitHalf_Polar_HFn <- rbind(
      ICC_SplitHalf_Polar_HFn,
      data.frame(
        Event = event,
        ICC_A1 = icc_value,
        Lower = lower_bound,
        Upper = upper_bound
      )
    )
  }
}

View(ICC_SplitHalf_Polar_HFn)

```


#### For Polar HR

```{r compute-epochs-icc-polar-hr}

#Compute ICC 
ICC_SplitHalf_Polar_HR <- data.frame()
events <- unique(Polar_HR_Epoch_Summary$Event_Phase)

for(event in events) {

  # Subset event
  df_event <- subset(
    Polar_HR_Epoch_Summary,
    Event_Phase == event
  )

  # Create odd/even split
  df_event <- df_event %>%

    mutate(
      Half = ifelse(
        Epoch %% 2 == 0,
        "Even",
        "Odd"
      )
    ) %>%

    # Mean lnRMSSD per participant × half
    group_by(ID, Half) %>%

    summarise(
      HR = mean(
        Mean_HR,
        na.rm = TRUE
      ),
      .groups = "drop"
    ) %>%

    # Wide format: Odd vs Even columns
    pivot_wider(
      names_from = Half,
      values_from = HR
    )

  # Keep complete cases
  df_event <- df_event[
    complete.cases(
      df_event$Odd,
      df_event$Even
    ),
  ]

  # Compute ICC if enough participants
  if(nrow(df_event) > 2) {

    icc <- ICC(
      df_event[, c("Odd", "Even")]
    )

    # Extract ICC(A,1)
    icc_row <- icc$results[
  "Single_raters_absolute",
  ]
 icc_value <- round(icc_row["ICC"],2)
    lower_bound <- round(icc_row["lower bound"],2)
    upper_bound <- round(icc_row["upper bound"],2)

    # Save result
    ICC_SplitHalf_Polar_HR <- rbind(
      ICC_SplitHalf_Polar_HR,
      data.frame(
        Event = event,
        ICC_A1 = icc_value, 
        Lower = lower_bound, 
        Upper = upper_bound 
      )
    )
  }
}

view(ICC_SplitHalf_Polar_HR)

```


## 4.2 Bland Altman Analysis

### 4.2.1 Preparation 

#### For RMSSD 

```{r compute-hr-bland-altman-stats}

# 0. Set theoretical / acceptable HR limits
theoretical_lower_rmssd <- -5.6
theoretical_upper_rmssd <- 5.6

# 1. Prepare paired HR data
RMSSD_BA <- Cardio_Comparison %>%
  filter(
    !is.na(RMSSD_Empatica),
    !is.na(RMSSD_Polar),
    !is.na(Event_Phase)
  ) %>%
  mutate(
    # Set facet order
    Event_Phase = factor(
      Event_Phase,
      levels = c("Baseline", "Trials", "Drone")
  )) %>%
  mutate(
    mean_RMSSD = (RMSSD_Empatica + RMSSD_Polar) / 2,
    diff_RMSSD = RMSSD_Empatica - RMSSD_Polar)


# 2. Bland-Altman analysis
RMSSD_BA_stats <- RMSSD_BA %>%
  group_by(Event_Phase) %>%
  summarise(
    n = n(),
    
    # Bland-Altman statistics
    bias = mean(diff_RMSSD),
    sd_difference = sd(diff_RMSSD),
    loa_lower = bias - 1.96 * sd_difference,
    loa_upper = bias + 1.96 * sd_difference,
    
    # Percentage inside theoretical boundaries
    n_inside = sum(
      diff_RMSSD >= theoretical_lower_rmssd &
      diff_RMSSD <= theoretical_upper_rmssd 
    ),
    
    pct_inside = 100 * n_inside / n,
    
    .groups = "drop"
  ) %>%
  mutate(
    pct_label = paste0(
      round(pct_inside, 0),
      "% of the data lies inside the boundaries."
    )
  )

RMSSD_BA_stats

```

#### For HFN 

```{r compute-hr-bland-altman-stats}

# 0. Set theoretical / acceptable HFn limits
theoretical_lower_hfn <- -0.04
theoretical_upper_hfn <-  0.04

# 1. Prepare paired HR data
HFn_BA <- Cardio_Comparison %>%
  filter(
    !is.na(HRV_HFn_Empatica),
    !is.na(HRV_HFn_Polar),
    !is.na(Event_Phase)
  ) %>%
  mutate(
    # Set facet order
    Event_Phase = factor(
      Event_Phase,
      levels = c("Baseline", "Trials", "Drone")
  )) %>%
  mutate(
    mean_HFn = (HRV_HFn_Empatica + HRV_HFn_Polar) / 2,
    diff_HFn = HRV_HFn_Empatica - HRV_HFn_Polar)


# 2. Bland-Altman analysis
HFn_BA_stats <- HFn_BA %>%
  group_by(Event_Phase) %>%
  summarise(
    n = n(),
    
    # Bland-Altman statistics
    bias = mean(diff_HFn),
    sd_difference = sd(diff_HFn),
    loa_lower = bias - 1.96 * sd_difference,
    loa_upper = bias + 1.96 * sd_difference,
    
    # Percentage inside theoretical boundaries
    n_inside = sum(
      diff_HFn >= theoretical_lower_hfn &
        diff_HFn <= theoretical_upper_hfn,
      na.rm = TRUE),
  
    pct_inside = 100 * n_inside / n,
    
    .groups = "drop"
  ) %>%
  mutate(
    pct_label = paste0(
      round(pct_inside, 0),
      "% of the data lies inside the boundaries."
    )
  )

HFn_BA_stats

```

#### For HR 

```{r compute-hr-bland-altman-stats}

# 0. Set theoretical / acceptable HR limits
theoretical_lower_hr <- -5
theoretical_upper_hr <-  5

# 1. Prepare paired HR data
HR_BA <- Cardio_Comparison %>%
  filter(
    !is.na(HR_Empatica),
    !is.na(HR_Polar),
    !is.na(Event_Phase)
  ) %>%
  mutate(
    # Set facet order
    Event_Phase = factor(
      Event_Phase,
      levels = c("Baseline",  "Trials",  "Drone")
  )) %>%
  mutate(
    mean_HR = (HR_Empatica + HR_Polar) / 2,
    diff_HR = HR_Empatica - HR_Polar)


# 2. Bland-Altman analysis
HR_BA_stats <- HR_BA %>%
  group_by(Event_Phase) %>%
  summarise(
    n = n(),
    
    # Bland-Altman statistics
    bias = mean(diff_HR),
    sd_difference = sd(diff_HR),
    loa_lower = bias - 1.96 * sd_difference,
    loa_upper = bias + 1.96 * sd_difference,
    
    # Percentage inside theoretical boundaries
    n_inside = sum(
      diff_HR >= theoretical_lower_hr &
      diff_HR <= theoretical_upper_hr
    ),
    
    pct_inside = 100 * n_inside / n,
    
    .groups = "drop"
  ) %>%
  mutate(
    pct_label = paste0(
      round(pct_inside, 0),
      "% of the data lies inside the boundaries."
    )
  )

HR_BA_stats

```

### 4.2.2 Assumptions 

#### For RMSSD

```{r bland-altman-assumptions-rmssd}

## Histogram
ggplot(RMSSD_BA, aes(x=mean_RMSSD)) + geom_histogram()+ facet_wrap(~Event_Phase) + theme_apa()


#Check Normality 
RMSSD_normality <- RMSSD_BA %>%
  group_by(Event_Phase) %>%
  summarise(
    shapiro_p = shapiro.test(diff_RMSSD)$p.value,
    .groups = "drop"
  ) %>%
  mutate(
    facet_label_rmssd = if_else(
      shapiro_p < .05,
      paste0(as.character(Event_Phase), "\u2020"),
      as.character(Event_Phase)
    )
  )

#Create Vector to use for labeling plot
facet_labels_rmssd <- setNames(
  RMSSD_normality$facet_label_rmssd,
  RMSSD_normality$Event_Phase
)

```

#### For HFn 

```{r bland-altman-assumptions-frequency}

## Histogram
ggplot(HFn_BA, aes(x=mean_HFn)) + geom_histogram()+ facet_wrap(~Event_Phase) + theme_apa()


#Check Normality 
HFn_normality <- HFn_BA %>%
  group_by(Event_Phase) %>%
  summarise(
    shapiro_p = shapiro.test(diff_HFn)$p.value,
    .groups = "drop"
  ) %>%
  mutate(
    facet_label_hfn = if_else(
      shapiro_p < .05,
      paste0(as.character(Event_Phase), "\u2020"),
      as.character(Event_Phase)
    )
  )

#Create Vector to use for labeling plot
facet_labels_hfn <- setNames(
  HFn_normality$facet_label_hfn,
  HFn_normality$Event_Phase
)

```

#### For HR

```{r bland-altman-assumptions-hr}

## Histogram
ggplot(HR_BA, aes(x=mean_HR)) + geom_histogram()+ facet_wrap(~Event_Phase) + theme_apa()


#Check Normality 
HR_normality <- HR_BA %>%
  group_by(Event_Phase) %>%
  summarise(
    shapiro_p = shapiro.test(diff_HR)$p.value,
    .groups = "drop"
  ) %>%
  mutate(
    facet_label_hr = ifelse(
      shapiro_p < .05,
      paste0(Event_Phase, " †"),
      Event_Phase
    )
  )

HR_normality

#Create Vector to use for labeling plot
facet_labels_hr <- setNames(
  HR_normality$facet_label_hr,
  HR_normality$Event_Phase
)

```


### 4.3 Visualization of Bland-Altman Plots

#### For RMSSD

```{r bland-altman-plots-rmssd}

# Bland-Altman plot by Event_Grouped
RMSSD_BA_plot <- ggplot(
  RMSSD_BA,
  aes(x = mean_RMSSD, y = diff_RMSSD)
) +
  
  # Observations
  geom_point(
    alpha = 0.55,
    size = 2
  ) +
  
    # Mean bias -- PURPLE
  geom_hline(
    data = RMSSD_BA_stats,
    aes(yintercept = bias),
    inherit.aes = FALSE,
    color = "#9b59d0",
    linewidth = 0.8
  ) +
  

  # Statistical 95% LoA -- RED
  geom_hline(
    data = RMSSD_BA_stats,
    aes(yintercept = loa_upper),
    inherit.aes = FALSE,
    color = "#d9534f",
    linewidth = 0.8
  ) +
  
  geom_hline(
    data = RMSSD_BA_stats,
    aes(yintercept = loa_lower),
    inherit.aes = FALSE,
    color = "#d9534f",
    linewidth = 0.8
  ) +
  
    # Theoretical / acceptable limits -- GREEN
  geom_hline(
    yintercept = theoretical_upper_rmssd,
    color = "#3fae5a",
    linewidth = 0.9
  ) +
  
  geom_hline(
    yintercept = theoretical_lower_rmssd,
    color = "#3fae5a",
    linewidth = 0.9
  ) +
  
    # Percentage inside theoretical boundaries
  geom_label(
    data = RMSSD_BA_stats,
    aes(
      x = -Inf,
      y = -Inf,
      label = pct_label
    ),
    inherit.aes = FALSE,
    
    hjust = -0.05,
    vjust = -0.5,
    
    fill = "white", 
    color = "#3fae5a",      
    text.colour = "black",
    size = 5, 
    
    label.size = 0.25,
    label.padding = unit(0.15, "lines"),

  ) +
  
     # Separate panel for each event
facet_wrap(
  ~ Event_Phase,
  ncol = 3,
  labeller = as_labeller(facet_labels_rmssd)
) +

# Labels & Theme
labs(
  title = "RMSSD Agreement by Experimental Event",
  subtitle = "† Differences are not normally distributed (Shapiro-Wilk p < .05)",
  x = "Mean RMSSD of Empatica and Polar (ms)",
  y = "Differemce (Empatica − Polar) (ms)"
) +
  
   theme_apa() +
  theme(
    
    plot.title = element_text(
      face = "bold",
      size = 15
    ),
    
    strip.text = element_text(
      face = "bold",
      size = 13
    ),
    
    axis.title = element_text(
      face = "bold", 
      size = 13
    ),
    
    axis.text = element_text(
      size = 13
    )
  )


RMSSD_BA_plot

```

#### For HFn

```{r bland-altman-plots-hfn}

# Bland-Altman plot by Event_Grouped
HFn_BA_plot <- ggplot(
  HFn_BA,
  aes(x = mean_HFn, y = diff_HFn)
) +
  
  # Observations
  geom_point(
    alpha = 0.55,
    size = 2
  ) +
  
    # Mean bias -- PURPLE
  geom_hline(
    data = HFn_BA_stats,
    aes(yintercept = bias),
    inherit.aes = FALSE,
    color = "#9b59d0",
    linewidth = 0.8
  ) +
  

  # Statistical 95% LoA -- RED
  geom_hline(
    data = HFn_BA_stats,
    aes(yintercept = loa_upper),
    inherit.aes = FALSE,
    color = "#d9534f",
    linewidth = 0.8
  ) +
  
  geom_hline(
    data = HFn_BA_stats,
    aes(yintercept = loa_lower),
    inherit.aes = FALSE,
    color = "#d9534f",
    linewidth = 0.8
  ) +
  
    # Theoretical / acceptable limits -- GREEN
  geom_hline(
    yintercept = theoretical_upper_hfn,
    color = "#3fae5a",
    linewidth = 0.9
  ) +
  
  geom_hline(
    yintercept = theoretical_lower_hfn,
    color = "#3fae5a",
    linewidth = 0.9
  ) +
  
    # Percentage inside theoretical boundaries
  geom_label(
    data = HFn_BA_stats,
    aes(
      x = -Inf,
      y = -Inf,
      label = pct_label
    ),
    inherit.aes = FALSE,
    
    hjust = -0.05,
    vjust = -0.5,
    
    fill = "white", 
    color = "#3fae5a",      
    text.colour = "black",  
        size = 5, 
    
    label.size = 0.25,
    label.padding = unit(0.15, "lines"),
  
  ) +
  
   # Separate panel for each event
facet_wrap(
  ~ Event_Phase,scales = "free", 
  ncol = 3,
  labeller = as_labeller(facet_labels_hfn)
) +

# Labels & Theme
labs(
  title = "HFn Agreement by Experimental Event",
  subtitle = "† Differences are not normally distributed (Shapiro-Wilk p < .05)",
  x = "Mean HFn of Empatica and Polar (n.u.)",
  y = "Difference (Empatica − Polar) (n.u.)"
) +
  
   theme_apa() +
  theme(
    
    plot.title = element_text(
      face = "bold",
      size = 15
    ),
    
    strip.text = element_text(
      face = "bold",
      size = 13
    ),
    
    axis.title = element_text(
      face = "bold", 
      size = 13
    ),
    
    axis.text = element_text(
      size = 13
    )
  )


HFn_BA_plot

#Explore potential Decreasing Overestimation with lower HFn values
ggplot(HFn_BA, aes(x=mean_HFn, y=diff_HFn)) + geom_point() + geom_smooth(method="lm") + theme_apa()
cor.test(HFn_BA$mean_HFn, HFn_BA$diff_HFn, method="spearman")
```

#### For HR 

```{r visualization-bland-altman-ht}

# Bland-Altman plot by Event_Grouped
HR_BA_plot <- ggplot(
  HR_BA,
  aes(x = mean_HR, y = diff_HR)
) +
  
  # Observations
  geom_point(
    alpha = 0.55,
    size = 2
  ) +
  
    # Mean bias -- PURPLE
  geom_hline(
    data = HR_BA_stats,
    aes(yintercept = bias),
    inherit.aes = FALSE,
    color = "#9b59d0",
    linewidth = 0.8
  ) +
  

  # Statistical 95% LoA -- RED
  geom_hline(
    data = HR_BA_stats,
    aes(yintercept = loa_upper),
    inherit.aes = FALSE,
    color = "#d9534f",
    linewidth = 0.8
  ) +
  
  geom_hline(
    data = HR_BA_stats,
    aes(yintercept = loa_lower),
    inherit.aes = FALSE,
    color = "#d9534f",
    linewidth = 0.8
  ) +
  
    # Theoretical / acceptable limits -- GREEN
  geom_hline(
    yintercept = theoretical_upper_hr,
    color = "#3fae5a",
    linewidth = 0.9
  ) +
  
  geom_hline(
    yintercept = theoretical_lower_hr,
    color = "#3fae5a",
    linewidth = 0.9
  ) +
  
    # Percentage inside theoretical boundaries
  geom_label(
    data = HR_BA_stats,
    aes(
      x = -Inf,
      y = -Inf,
      label = pct_label
    ),
    inherit.aes = FALSE,
    
    hjust = -0.05,
    vjust = -0.5,
    
    fill = "white", 
    color = "#3fae5a",      
    text.colour = "black",
    size = 5, 
    
    label.size = 0.25,
    label.padding = unit(0.15, "lines"),
  ) +
  
    # Separate panel for each event
  facet_wrap(
  ~ Event_Phase, scales = "free", 
  ncol = 3,
  labeller = as_labeller(facet_labels_hr)
) +

# Labels & Theme
labs(
  title = "HR Agreement by Experimental Event",
  subtitle = "† Differences are not normally distributed (Shapiro-Wilk p < .05)",
  x = "Mean HR of Empatica and Polar (bpm)",
  y = "Difference (Empatica - Polar) (bpm)"
) +
  
  theme_minimal(base_size = 11) +
  theme(
    
    plot.title = element_text(
      face = "bold",
      size = 15
    ),
    
    strip.text = element_text(
      face = "bold",
      size = 13
    ),
    
    axis.title = element_text(
      face = "bold", 
      size = 13
    ),
    
    axis.text = element_text(
      size = 13
    )
  )


HR_BA_plot

#Explore potential increase in underestimation with higher HR values
ggplot(HR_BA, aes(x=mean_HR, y=diff_HR)) + geom_point() + geom_smooth(method="lm") + theme_apa()
cor.test(HR_BA$mean_HR, HR_BA$diff_HR, method="spearman")


```

### Save Plots 

```{r save-bland-altman-plots}

#RMSSD
ggsave(
  "RMSSD_BA_plot.png",
  RMSSD_BA_plot,
  width = 12,
  height = 8,
  units = "in",
  dpi = 600
)

#HFn
ggsave(
  "HFn_BA_plot.png",
  HFn_BA_plot,
  width = 12,
  height = 8,
  units = "in",
  dpi = 600
)

#HR
ggsave(
  "HR_BA_plot.png",
  HR_BA_plot,
  width = 12,
  height = 8,
  units = "in",
  dpi = 600
)

```


## 4.4 Agreement between Sensors: ICC (A1) 

### For RMSSD

```{r agreemment-between-sensors-rmssd}

#Compute ICC 
icc_results_rmssd <- Cardio_Comparison %>%
  
  filter(
    complete.cases(
      lnRMSSD_Polar,
      lnRMSSD_Empatica
    )
  ) %>%
  
  group_by(Event_Phase) %>%
  
  group_modify(~{
    
    if (nrow(.x) <= 2) {
      return(tibble())
    }
    
    icc <- ICC(
      .x %>%
        select(
          lnRMSSD_Polar,
          lnRMSSD_Empatica
        )
    )
    
    data.frame(
      Type = rownames(icc$results),
      icc$results,
      row.names = NULL
    )
  }) %>%
  
  ungroup()

#View Results
icc_results_rmssd <- icc_results_rmssd %>% 
  filter(Type == "Single_raters_absolute") %>% dplyr:: select(c(Event_Phase, Type, ICC, lower.bound, upper.bound, F,df1,df2,p, type ))


```

### For HFn

```{r agreemment-between-sensors-frequency}

#Compute ICC 
icc_results_hfn <- Cardio_Comparison %>%
  
  filter(
    complete.cases(
      HRV_HFn_Polar,
      HRV_HFn_Empatica
    )
  ) %>%
  
  group_by(Event_Phase) %>%
  
  group_modify(~{
    
    if (nrow(.x) <= 2) {
      return(tibble())
    }
    
    icc <- ICC(
      .x %>%
        select(
          HRV_HFn_Polar,
          HRV_HFn_Empatica
        )
    )
    
    data.frame(
      Type = rownames(icc$results),
      icc$results,
      row.names = NULL
    )
  }) %>%
  
  ungroup()

#View Results
icc_results_hfn <- icc_results_hfn %>% 
  filter(Type == "Single_raters_absolute") %>% dplyr:: select(c(Event_Phase, Type, ICC, lower.bound, upper.bound, F,df1,df2,p, type ))

icc_results_hfn

```

### For HR

```{r agreemment-between-sensors-hr}

#Compute ICC 
icc_results_hr <- Cardio_Comparison %>%
  
  filter(
    complete.cases(
      HR_Polar,
      HR_Empatica
    )
  ) %>%
  
  group_by(Event_Phase) %>%
  
  group_modify(~{
    
    if (nrow(.x) <= 2) {
      return(tibble())
    }
    
    icc <- ICC(
      .x %>%
        select(
          HR_Polar,
          HR_Empatica
        )
    )
    
    data.frame(
      Type = rownames(icc$results),
      icc$results,
      row.names = NULL
    )
  }) %>%
  
  ungroup()

#View Results
icc_results_hr <- icc_results_hr %>% 
  filter(Type == "Single_raters_absolute") %>% dplyr:: select(c(Event_Phase, Type, ICC,lower.bound, upper.bound, F,df1,df2,p, type ))


```

## 4.5 Correlations between Sensors

### RMSSD

```{r correlations-between-sensors-rmssd}

# Get event names
events <- unique(Cardio_Comparison$Event_Phase)

# Empty table for correlations
cor_results <- data.frame()

# Empty list for plots
plot_list <- list()

for(event in events) {

  # Subset data
  df_event <- subset(
    Cardio_Comparison,
    Event_Phase == event
  )

  # Remove missing values
  df_event <- df_event[
    complete.cases(
      df_event$lnRMSSD_Polar,
      df_event$lnRMSSD_Empatica
    ),
  ]

  # Only continue if enough data
  if(nrow(df_event) > 2) {

    # Pearson correlation
    r <- cor(
      df_event$lnRMSSD_Polar,
      df_event$lnRMSSD_Empatica,
      method = "spearman"
    )

    # p-value
    cor_test <- cor.test(
      df_event$lnRMSSD_Polar,
      df_event$lnRMSSD_Empatica,
      method = "spearman"
    )

    # Save results
    cor_results <- rbind(
      cor_results,
      data.frame(
        Event_Grouped = event,
        Pearson_r = r,
        p_value = cor_test$p.value,
        N = nrow(df_event)
      )
    )

    # Create plot
    p <- ggplot(
      df_event,
      aes(
        x = lnRMSSD_Polar,
        y = lnRMSSD_Empatica
      )
    ) +

      geom_point(size = 2) +

      geom_smooth(
        method = "lm",
        se = FALSE
      ) +

      labs(
        title = paste0(
          event,
          " (r = ", round(r, 2),", p = ", round(cor_test$p.value,3),")"
        ),
        x = "Polar lnRMSSD",
        y = "Empatica lnRMSSD"
      ) +

      theme_minimal()

    # Store plot
    plot_list[[event]] <- p
  }
}

# Show correlation results
print(cor_results)
wrap_plots(plot_list, ncol = 4)

```

### HFn

```{r correlations-between-sensors-hfn}

# Get event names
events <- unique(Cardio_Comparison$Event_Phase)

# Empty table for correlations
cor_results <- data.frame()

# Empty list for plots
plot_list <- list()

for(event in events) {

  # Subset data
  df_event <- subset(
    Cardio_Comparison,
    Event_Phase == event
  )

  # Remove missing values
  df_event <- df_event[
    complete.cases(
      df_event$HRV_HFn_Polar,
      df_event$HRV_HFn_Empatica
    ),
  ]

  # Only continue if enough data
  if(nrow(df_event) > 2) {

    # Pearson correlation
    r <- cor(
      df_event$HRV_HFn_Polar,
      df_event$HRV_HFn_Empatica,
      method = "spearman"
    )

    # p-value
    cor_test <- cor.test(
      df_event$HRV_HFn_Polar,
      df_event$HRV_HFn_Empatica,
      method = "spearman"
    )

    # Save results
    cor_results <- rbind(
      cor_results,
      data.frame(
        Event_Grouped = event,
        Pearson_r = r,
        p_value = cor_test$p.value,
        N = nrow(df_event)
      )
    )

    # Create plot
    p <- ggplot(
      df_event,
      aes(
        x = HRV_HF_Polar,
        y = HRV_HF_Empatica
      )
    ) +

      geom_point(size = 2) +

      geom_smooth(
        method = "lm",
        se = FALSE
      ) +

      labs(
        title = paste0(
          event,
          " (r = ", round(r, 2),", p = ", round(cor_test$p.value,3),")"
        ),
        x = "Polar HRV_HFn",
        y = "Empatica HRV_HFn"
      ) +

      theme_minimal()

    # Store plot
    plot_list[[event]] <- p
  }
}

# Show correlations
print(cor_results)
wrap_plots(plot_list, ncol = 4)


```

### HR

```{r correlations-between-sensors-hr}

# Get event names
events <- unique(Cardio_Comparison$Event_Phase)

# Empty table for correlations
cor_results <- data.frame()

# Empty list for plots
plot_list <- list()

for(event in events) {

  # Subset data
  df_event <- subset(
    Cardio_Comparison,
    Event_Phase  == event
  )

  # Remove missing values
  df_event <- df_event[
    complete.cases(
      df_event$HR_Polar,
      df_event$HR_Empatica
    ),
  ]

  # Only continue if enough data
  if(nrow(df_event) > 2) {

    # Pearson correlation
    r <- cor(
      df_event$HR_Polar,
      df_event$HR_Empatica,
      method = "spearman"
    )

    # p-value
    cor_test <- cor.test(
      df_event$HR_Polar,
      df_event$HR_Empatica,
      method = "spearman"
    )

    # Save results
    cor_results <- rbind(
      cor_results,
      data.frame(
        Event_Grouped = event,
        Pearson_r = r,
        p_value = cor_test$p.value,
        N = nrow(df_event)
      )
    )

    # Create plot
    p <- ggplot(
      df_event,
      aes(
        x = HR_Polar,
        y = HR_Empatica
      )
    ) +

      geom_point(size = 2) +

      geom_smooth(
        method = "lm",
        se = FALSE
      ) +

      labs(
        title = paste0(
          event,
        " (r = ", round(r, 2),", p = ", round(cor_test$p.value,3),")"
        ),
        x = "Polar HR",
        y = "Empatica HR"
      ) +

      theme_minimal()

    # Store plot
    plot_list[[event]] <- p
  }
}

# Show correlation table
print(cor_results)
wrap_plots(plot_list, ncol = 4)

```


# 5. Event Level Analysis

## 5.1 Prepare: Long dataset 

```{r prep-long-dataset}

#Create Long Dataset
Cardio_Comparison_long <- Cardio_Comparison %>%
  pivot_longer(
    cols = matches("_(Empatica|Polar)$"),
    names_to = c(".value", "Device"),
    names_pattern = "(.*)_(Empatica|Polar)$"
  ) 

#Only look at Baseline & Fatigue Induction (AF & PF)
Cardio_Comparison_long_F <- Cardio_Comparison_long %>% 
  filter (!Event_Phase %in% c("Drone")) %>% 
  filter(Group != "Control")


#Z-Score
Cardio_Comparison_long_F <- Cardio_Comparison_long_F %>%
  group_by(Device) %>%
  mutate(lnRMSSD_z = as.numeric(scale(lnRMSSD)), 
         HFn_z = as.numeric(scale(HRV_HFn)), 
         HR_z = as.numeric(scale(HR))) %>%
  ungroup()

```

## 5.2 Visual Inspection: Fatigue Type 

```{r difference-fatigue}

#lnRMSSD
Cardio_Comparison_long_F %>%
ggplot(aes(x=Event_Phase, y=lnRMSSD, fill=Group)) + geom_boxplot() + facet_grid(~Device) +theme_minimal()


#HRV HF
Cardio_Comparison_long_F %>%
ggplot(aes(x=Event_Phase, y=HRV_HF, fill=Group)) + geom_boxplot() + facet_grid(~Device) +theme_minimal()
  

#HR Quo
Cardio_Comparison_long_F %>% 
ggplot(aes(x=Event_Phase, y=HR, fill=Group)) + geom_boxplot() + facet_grid(~Device) +theme_minimal()

```

## 5.3 Linear Mixed-Model

To test, if both devices (Polar and Empatica) lead to the same  conclusions about the experimental manipulation, we will utilize a linear mixed model. 
We will only be looking at the potential change from baseline to fatigue condition (introduced by the fatigue manipulation). 

For each CVA metric, we will specify the following formula: 

-> CVA_Outcome ~ Group * Event_Phase * Device + Random_Intercept

#### LnRMSSD Model 

```{r mixed-model-rmssd}

#Build Model 
mod_rmssd <- lmer(lnRMSSD_z ~ Group * Event_Phase * Device +    (1 | ID), data= Cardio_Comparison_long_F)
summary(mod_rmssd)
anova(mod_rmssd)

#Visualize the EMM (separated by Group)
plot_emm_rmssd <- emmip(
  mod_rmssd,
  Device ~ Event_Phase | Group,
  CIs = TRUE) + theme_minimal() + 
  labs(
    title = paste("Estimated Marginal Means"),
    subtitle = paste("lnRMSSD (z-scored)"),
    x = "Event",
    y = "lnRMSSD"
  )  + theme_apa() +
  theme(
    plot.title = element_text(
      face = "bold",
      size = 14
    ),
     strip.text = element_text(
      face = "bold",
      size = 14
    ),
    axis.title = element_text(
      face = "bold",
      size = 14
  ), 
    
    axis.text = element_text(
      size = 14), 
    legend.text = element_text(size = 12),
  legend.title = element_text(size = 12))


# Estimated marginal means of Phase
# separately for each Device × Group combination
emm_rmssd <- emmeans(
  mod_rmssd,
  ~ Event_Phase | Device * Group)
pairs(
  emm_rmssd,
    reverse = TRUE,
  adjust = "BH")
```


#### HFn Model 

```{r mixed-model-rmssd}

#Build Model 
mod_hfn <- lmer(HFn_z ~ Group * Event_Phase * Device +    (1 | ID), data= Cardio_Comparison_long_F)
summary(mod_hfn)
anova(mod_hfn)


#Visualize the EMM (separated by Group)
plot_emm_hfn <- emmip(
  mod_hfn,
  Device ~ Event_Phase | Group,
  CIs = TRUE) + theme_minimal() +
  labs(
    title = paste("Estimated Marginal Means"),
    subtitle = paste("Normalized High Frequency (z-scored)"),
    x = "Event",
    y = "HFn"
  ) + theme_apa() +
  theme(
    plot.title = element_text(
      face = "bold",
      size = 14
    ),
     strip.text = element_text(
      face = "bold",
      size = 14
    ),
    axis.title = element_text(
      face = "bold",
      size = 14
  ), 
    
    axis.text = element_text(
      size = 14), 
    legend.text = element_text(size = 12),
  legend.title = element_text(size = 12))

# Estimated marginal means of Phase
emm_hfn <- emmeans(
  mod_hfn,
  ~ Event_Phase | Device * Group)
pairs(
  emm_hfn,
   reverse = TRUE,
  adjust = "BH")
```

#### HR 

```{r mixed-model-hr}

#Build Model 
mod_hr <- lmer(HR_z ~ Group * Event_Phase * Device +    (1 | ID), data= Cardio_Comparison_long_F)
summary(mod_hr)
anova(mod_hr)


#Visualize the EMM (separated by Group)
plot_emm_hr <- emmip(
  mod_hr,
  Device ~ Event_Phase | Group,
  CIs = TRUE) + theme_minimal() +
  labs(
    title = paste("Estimated Marginal Means"),
    subtitle = paste("Heart Rate (z-scored)"),
    x = "Event",
    y = "HR"
  ) + theme_apa() +
    theme(
    plot.title = element_text(
      face = "bold",
      size = 14
    ),
     strip.text = element_text(
      face = "bold",
      size = 14
    ),
    axis.title = element_text(
      face = "bold",
      size = 14
  ), 
    
    axis.text = element_text(
      size = 14), 
    legend.text = element_text(size = 12),
  legend.title = element_text(size = 12))

# Estimated marginal means of Phase
emm_hr <- emmeans(
  mod_hr,
  ~ Event_Phase | Device * Group)
pairs(
  emm_hr,
    reverse = TRUE,
  adjust = "BH")
```

### Combined plot 

```{r put-emm-plots-in-one}

#Remove individual x-axis titles
plot_emm_rmssd <- plot_emm_rmssd + labs(x = NULL)
plot_emm_hfn   <- plot_emm_hfn   + labs(x = NULL)

#Combine
combined_plot <- (
  plot_emm_rmssd /
  plot_emm_hfn /
  plot_emm_hr
) +
  plot_layout(
    guides = "collect",
    heights = c(1, 1, 1)
  ) +
  plot_annotation(
    tag_levels = "A",
    theme = theme(
      plot.tag = element_text(face = "bold", size = 12)
    )
  )+ coord_cartesian(ylim = c(-2, 2))

combined_plot

# Save larger
ggsave(
  "Combined_EMM_plot.png",
  combined_plot,
  width = 12,
  height = 8,
  units = "in",
  dpi = 600
)

```


## 5.4 Check Assumptions of Models

### For lnRMSSD

```{r assumptions-lnrmssd-model}

#Check Assumptions 

## Plots
check_rmssd <- check_model(mod_rmssd)
windows()
plot(check_rmssd)


## Functions
performance::check_normality(mod_rmssd)
performance::check_heteroscedasticity(mod_rmssd)
performance::check_singularity(mod_rmssd)
r2(mod_rmssd)

## Normality of Random Intercept
ri_rmssd <- ranef(mod_rmssd)$ID
qqnorm(ri_rmssd[, "(Intercept)"])
qqline(ri_rmssd[, "(Intercept)"], col = "red")
shapiro.test(ri_rmssd[, "(Intercept)"])

icc(mod_rmssd)


```

### For HFn 

```{r assumptions-hfn-model}

#Check Assumptions 

## Plots
check_hfn <- check_model(mod_hfn)
windows()
plot(check_hfn)
dev.off()


## Functions
performance::check_normality(mod_hfn)
performance::check_heteroscedasticity(mod_hfn)
performance::check_singularity(mod_hfn)
r2(mod_hfn)


## Normality of Random Intercept
ri_hfn <- ranef(mod_hfn)$ID
qqnorm(ri_hfn[, "(Intercept)"])
qqline(ri_hfn[, "(Intercept)"], col = "red")
shapiro.test(ri_hfn[, "(Intercept)"])


icc(mod_hfn)

```

### For HR 

```{r assumptions-hr-model}

#Check Assumptions 

## Plots
check_hr <- check_model(mod_hr)
windows()
plot(check_hr)
dev.off()

## Functions
performance::check_normality(mod_hr)
performance::check_heteroscedasticity(mod_hr)
performance::check_singularity(mod_hr)
r2(mod_hr)


## Normality of Random Intercept
ri_hr <- ranef(mod_hr)$ID
qqnorm(ri_hr[, "(Intercept)"])
qqline(ri_hr[, "(Intercept)"], col = "red")
shapiro.test(ri_hr[, "(Intercept)"])


plot(fitted(mod_hr), resid(mod_hr))
abline(h = 0, col = "red")

icc(mod_hr) 

```

