---
title: "Fatigue Physio Preprocessing"
date: "`r Sys.Date()`"
output: html_document
---

This is the preprocessing of cardiovascular activity data of the study \_\_\_\_\_\_\_. For further information on the study, please refer to the corresponding paper.

---

# 0. Packages

a)  Download Data Wrangling Packages

```{r data-wrangling-packages,include=FALSE	}

library(readxl) #To import the data
library(tidyverse) #To prepare the data adequately
library(magrittr) #To ensure seamless workflow
library(writexl) #To export data
library(purrr) #programming tools 
library(stringr) #wrappers
```

b)  Download Statistics & Analysis Packages

```{r statistics-packages, include=FALSE	}

library (Hmisc) #Harrell Miscellaneous
library(jtools) #analysis and presentation of data
library(knitr) #reporting 
```

c)  Download TimeSeries-Related Packages

```{r timeseries-packages, include=FALSE	}

library(imputeTS) #timeseries imputation
library(slider) #sliding window function
library(pracma) #for imputations (math function package)
library(hms) #for pretty time of day
library(lubridate) #parse & manipulate dates 
```

# 1. Import Data

## 1.1 Import TimeStamps & Performance

```{r import-timestamps}

#Import Data 
Timestamps <- read_excel("../Supplementary_Inputs/Timestamps_Final.xlsx", 
    col_types = c("text", "text", "numeric", 
        "text", "date", "date", "date", "date", 
        "date", "date", "date", "numeric", 
        "text", "text", "text", "text", "text", 
        "text", "text"))


#Quick-Timestamp Preprocess: 
time_cols <- c(
  "Start_Baseline", "End_Baseline",
  "End_Trail1", "End_Trail2", "End_Trail3",
  "Start_Drone", "End_Drone"
)

Timestamps <- Timestamps %>%
  mutate(across(all_of(time_cols),
                ~ format(.x, "%H:%M:%S")))

```

## 1.2. Import Survey Data

```{r import-self-reports}

#Data from after Task
Fatigue_Task <- read_excel("../Supplementary_Inputs/Fatigue_1.xlsx", 
    col_types = c("text", "text", "text", 
        "text", "text", "date", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text"))
Fatigue_Task <- Fatigue_Task[-1, ]

```

## 1.3 Import Empatica Data

```{r import-empatica-data}

#RR-Intervall Data 
Empatica_RR <- read_csv("CVA_Empatica/systolic_peaks_all_participants.csv")
Empatica_RR <- Empatica_RR %>% rename(ID = participant_id, Local_Time = local_time_sec)
Empatica_RR %>% distinct(ID) %>% print(n=50)

#Delete Empatica HR Data w/o Self-Report (No Participation in Full Study)
Empatica_RR <- Empatica_RR %>% 
  filter(!ID %in% c("OITC40", "OITC45", "OITC46")) #Ages: 24,26,30, Gender: F,M,M

#Delete Participant, which disregarded Instruction for n-Back Task
Empatica_RR <- Empatica_RR %>% 
  filter(!ID %in% c("OITC04"))

```

## 1.4 Import Polar Data: RR

```{r import-polar-data-rr}

#Set Working Directory
setwd("./CVA_Reference")

# List RR files that start with "ID" followed by digits
rrfiles <- list.files(pattern = "_RR", full.names = TRUE)

#Load all files into a list
rr_list <- lapply(rrfiles, read.csv)
names(rr_list) <- tools::file_path_sans_ext(basename(rrfiles)) #adopt names

```

## 1.5 Import Polar Data: HR
Needed for Meta-Data of the Polar recording.  

```{r import-polar-data-hr}

setwd("./CVA_Reference")
#List all HR-Files 
hrfiles <- list.files( pattern="_HR",full.names=TRUE)


#Load all files into a list
hr_list <- lapply(hrfiles, read.csv)
names(hr_list) <- tools::file_path_sans_ext(basename(hrfiles)) #adopt name

```

# 2. Preprocess: Clean Self-Report Data 

```{r clean-selfreport-data}

#Rename Columns
Fatigue_Task <- Fatigue_Task %>% rename(Session = IN06_04, Color = IN06_02, ID = IN06_01, Station = IN06_03, Start_Q1 = STARTED)
Fatigue_Task$ID[Fatigue_Task$ID == "01TC3"] <- "OITC03"

#Clean Identifier 

## Build Function
fix_id <- function(x) {
  # Extract numbers from ID
  nums <- gsub("\\D", "", x)
  
  # Rebuild as OITCXX
  paste0("OITC", sprintf("%02d", as.numeric(nums)))
}

## Apply
Fatigue_Task$ID <- fix_id(Fatigue_Task$ID)

```

# 3. Preprocess: Get Polar Data in the right Format

## 3.1 Participant Information  
Save Meta Data (Start-Time of recording) for each ID into a dataframe called "Participant_Info". 

```{r}

#1) Create Empty List
participant_info_list <- list()  # For metadata
cleaned_data_list <- list()      # For cleaned HR data

#2) Loop through each dataframe in the list
for (i in seq_along(hr_list)) {
  
  df <- hr_list[[i]]
  df_name <- names(hr_list)[i]
  
  ## a) Extract participant info from first row
  info <- data.frame(
    Date = df[1, 3],
    StartTime = df[1, 4],
    Duration = df[1, 5],
    Avg_HR = df[1, 7],
    File = df_name,
    stringsAsFactors = FALSE
  )
  
  participant_info_list[[i]] <- info
  
  ## b) Clean the rest of the data
  cleaned <- df[-(1:2), ]  # Remove first 2 rows
  cleaned <- cleaned %>%
    rename(Time = Sport, HR = Date) %>%
    select(Time, HR) #rename & select 
  cleaned$HR <- as.numeric(cleaned$HR)
  cleaned$File <- df_name  # Add file name to each row for ID
  
  cleaned_data_list[[df_name]] <- cleaned
}

#3) Combine all participant metadata into one dataframe
Participant_Info<- bind_rows(participant_info_list)
Polar_HR <- bind_rows(cleaned_data_list)


#4) Change File Name to ID 

## Extract ID from file name (e.g., "ID23_Whatever.csv" → "ID23")
extract_id <- function(filename) {
  sub("^((OITC[0-9]+)).*", "\\1", filename)
}

## Create ID
Participant_Info$ID <- extract_id(Participant_Info$File)
Participant_Info <- Participant_Info %>%
  select(ID, everything(), -c(File))

```

## 3.2 RR-Data 
Add a TimeStamp to the RR-Data from Polar. 
Add the ID (from the name of the dataframes), the real TimeStamp (from Start Time of Measurement) and a Seconds Counter to the RR-Intervals. 
Meta-Data is provided from "Participant_Info". 

```{r preproccess-RR-data}

#1) Add ID
updated_rr <- list() #create empty list document

## Loop through each dataframe in the list
for (df_name in names(rr_list)) {
  
  # Extract ID (e.g., "ID01") from the dataframe name
  id_match <- regmatches(df_name, regexec("^(OITC\\d+)_", df_name))[[1]]
  
  # If no ID found, skip
  if (length(id_match) < 2) {
    warning(paste("ID not found in:", df_name))
    next
  }
  
  id <- id_match[2]
  
  # Get the dataframe and add the ID column
  df <- rr_list[[df_name]] %>%
    mutate(ID = id) %>% select(c("ID", "offline", "duration")) %>% rename(RR = duration)
  
  # Store in updated list
  updated_rr[[df_name]] <- df
}

## Combine into one dataframe
Polar_RR <- bind_rows(updated_rr)

#2) Add TimeStamp

## Add StartTime to RR Data 
Polar_RR <-
  Polar_RR %>%
  left_join(Participant_Info %>% select(ID, StartTime), by = "ID")

##Start TimeStamp after first RR Intervall
Polar_RR <- Polar_RR %>%
  group_by(ID) %>%
  mutate(
    cum_duration_ms = cumsum(RR),
    
    Seconds = cum_duration_ms / 1000,
    
    Timestamp = hms::as_hms(
      as.numeric(lubridate::hms(StartTime)) + Seconds
    )
  ) %>% 
  ungroup()

##Adjust to start at 0.00?
RR_0 <- Polar_RR  %>%
  group_by(ID) %>%
  mutate(
    cum_duration_ms = cumsum(RR) - first(RR), # adjust to start at 0.00
    Seconds = cum_duration_ms / 1000,
  ) %>%
  ungroup()

```

# 4. Preprocess: Flagging & Filtering Data

Okay, so the following events need to be flagged in the RR-Data. This is the time-phases we want to keep & analyze. 

- Baseline Measurement (as Baseline): from Start_Baseline to End_Baseline

- Trial (n-Back & Video) Measurement:

  - in 1-Back & Dual-2-Back (as TrialX) End_Trial1 - 300, End_Trial2 - 300, End_Trial3 - 300

  - in Control (as Video) = Start_Q1 - 900

- Drone Measurement (as Drone): from Start_Drone to End_Drone

## 4.1 For Empatica Data

### 4.1.1 Event Flagging in Empatica Data

```{r event-flagging-in-empatica-data}

#Unify Data from Started Fatigue_Task to Timestamp 
Timestamps <- Timestamps %>%
  left_join(Fatigue_Task %>% dplyr::select(ID,Start_Q1), by = "ID")

Timestamps <- Timestamps %>% 
  mutate (Start_Q1 = format(Start_Q1, "%H:%M:%S")) 

#Event Flagging 

## Convert all timestamp columns to hms 
time_cols <- c( "Start_Baseline", "End_Baseline", "End_Trail1", "End_Trail2", "End_Trail3", "Start_Drone", "End_Drone", "Start_Q1" ) 
Timestamps <- Timestamps %>% mutate(across(all_of(time_cols), as_hms)) 

## Join + create flags
Empatica_RR <- Empatica_RR %>% 
  left_join(Timestamps, by = "ID") %>% 
  
  #Baseline 
  mutate( Event = case_when( 
    Local_Time >= Start_Baseline & 
    Local_Time <= End_Baseline ~ "Baseline", 
    
    #Trial 1 
    Group %in% c("1-Back", "Dual-2-Back") & !is.na(End_Trail1) &
      Local_Time >= (End_Trail1 - 300) & #
      Local_Time <= End_Trail1 ~ "Trial1", 
    
    # Trial 2 
    Group %in% c("1-Back", "Dual-2-Back") & !is.na(End_Trail2) & 
      Local_Time >= (End_Trail2 - 300) & Local_Time <= End_Trail2 ~ "Trial2",
    
    # Trial 3 
    Group %in% c("1-Back", "Dual-2-Back") & !is.na(End_Trail3) &
      Local_Time >= (End_Trail3 - 300) & Local_Time <= End_Trail3 ~ "Trial3", 
    
    # Control video condition 
    Group == "Control" & !is.na(Start_Q1) & Local_Time >= (Start_Q1 - 900) & Local_Time <= Start_Q1 ~ "Video", 
    
    # Drone 
    Local_Time >= Start_Drone & Local_Time <= End_Drone ~ "Drone", 
   
     # Everything else
    TRUE ~ "Irrelevant" ) )

```

### 4.1.2 Filtering

```{r filtering-in-empatica-data}

Empatica_RR_Relevant <- Empatica_RR %>%
  filter(Event != "Irrelevant") %>% dplyr::select(c(ID, Session, Color, Station, timestamp, Local_Time, rr_interval_ms, Event, Group, Type)) %>% rename (RR_Raw = rr_interval_ms)
Empatica_RR_Relevant %>% distinct(ID) %>% print(n=50)

```

## 4.2 For Polar Data 

### 4.2.1 Event Flagging in Polar Data: RR

```{r event-flagging-in-empatica-data}

#Event Flagging

## Join Timestamps + create flags
Polar_RR <- Polar_RR %>%
  left_join(Timestamps, by = "ID")

Polar_RR <- Polar_RR %>%
  
  mutate(
    
    ts  = as.numeric(Timestamp),
    
    sb  = as.numeric(Start_Baseline),
    eb  = as.numeric(End_Baseline),
    
    et1 = as.numeric(End_Trail1),
    et2 = as.numeric(End_Trail2),
    et3 = as.numeric(End_Trail3),
    
    sd  = as.numeric(Start_Drone),
    ed  = as.numeric(End_Drone),
    
    sq1 = as.numeric(Start_Q1),
    
    Event = case_when(
      
      # Baseline
      ts >= sb &
      ts <= eb ~ "Baseline",
      
      # Trial 1
      Group %in% c("1-Back", "Dual-2-Back") &
      ts >= (et1 - 300) &
      ts <= et1 ~ "Trial1",
      
      # Trial 2
      Group %in% c("1-Back", "Dual-2-Back") &
      ts >= (et2 - 300) &
      ts <= et2 ~ "Trial2",
      
      # Trial 3
      Group %in% c("1-Back", "Dual-2-Back") &
      ts >= (et3 - 300) &
      ts <= et3 ~ "Trial3",
      
      # Video
      Group == "Control" &
      ts >= (sq1 - 900) &
      ts <= sq1 ~ "Video",
      
      # Drone
      ts >= sd &
      ts <= ed ~ "Drone",
      
      TRUE ~ "Irrelevant"
    )
    
  ) %>%
  
  select(-ts, -sb, -eb, -et1, -et2, -et3, -sd, -ed, -sq1)
```

### 4.2.2 Filtering: RR

```{r filtering-in-empatica-data}

Polar_RR_Relevant <- Polar_RR %>%
  filter(Event != "Irrelevant") %>% dplyr::select(c(ID, Session, Color, Station, Timestamp, RR, Event, Group, Type, offline)) %>% rename (RR_Raw = RR)
Polar_RR_Relevant %>% distinct(ID) %>% print(n=50)

```

# 5. Artefact Detection & Imputation

Utilizing a two step-approach for detecting artefacts: 
1) Only keep absolute plausible values (Jafaar & Chung Xian, 2021)
2) Only keep plausible jumps (= moving local median) (Alcantara et al., 2020)

## 5.1 For Empatica Data 

### 5.1.1 Artifact Detection 

```{r artefact-detection-for-empatica-data}

#Artefact Detection 
Empatica_RR_Raw <- Empatica_RR_Relevant %>%  
  group_by(ID) %>%
  
  #Absolute Threshold
  mutate(
    Artifact_Flag = case_when(
      RR_Raw < 375 ~ "Artifact",
      RR_Raw > 1200 ~ "Artifact", 
      TRUE ~ "OK" )) %>% 
  
 #Relative Threshold
  mutate(
    Local_Median = slide_dbl(
      .x = RR_Raw,
      .before = 5,  
      .after = 5,   
      .f = ~ median(.x, na.rm = TRUE)),
    RR_Deviation = abs(RR_Raw - Local_Median),
    Artifact_Flag = case_when(
    Artifact_Flag == "Artifact" ~ "Artifact", 
    RR_Deviation > 250 ~ "Artifact", 
    TRUE ~ "OK")) %>%
  
  ungroup()

```

### 5.1.2 Artifact Imputation 

```{r artifcat-imputation-empatica}

#Change Flagged Artifacts into NAs
Empatica_RR_Flagged <- Empatica_RR_Raw %>% 
  mutate(RR_clean = if_else(Artifact_Flag == "Artifact", NA_real_, RR_Raw))


#Impute TS Function: NA_MA replaces NAs w/ moving average
Empatica_RR_Imputed <- Empatica_RR_Flagged %>%
  group_by(ID) %>%
  arrange(Local_Time) %>%
  group_modify(~ {
    .x$EWMA_RR <- na_ma(.x$RR_clean, k = 32, weighting = "exponential")
    .x
  }) %>%
  ungroup()

```

### 5.1.3 Imputation Visualization

```{r ewma-visualization-empatica}

# Select participant
selected_id <- "OITC28"

# Prepare plotting data
plot_data <- Empatica_RR_Imputed %>%
  
  filter(ID == selected_id) %>%
  filter (Event %in% c("Trial1", "Drone")) %>% 
  
  #group_by(Event) %>%
  
  arrange(Local_Time) %>%
  
  mutate(
    Seconds = row_number(),
    Imputed = is.na(RR_clean)
  ) %>%
  
  ungroup()

# Plot
ggplot(plot_data,
       aes(x = Seconds,
           y = EWMA_RR)) +
  
  # Main RR signal
  geom_line(color = "blue",
            linewidth = 0.4) +
  
  # Original artifacts
  geom_point(
    data = subset(plot_data, Imputed),
    aes(y = RR_Raw),
    color = "red",
    shape = 4,
    size = 2
  ) +
  
  # EWMA imputed values
  geom_point(
    data = subset(plot_data, Imputed),
    aes(y = EWMA_RR),
    color = "black",
    shape = 17,
    size = 2
  ) +
  
  # Connect raw artifact to imputed value
  geom_segment(
    data = subset(plot_data, Imputed),
    aes(
      xend = Seconds,
      y = RR_Raw,
      yend = EWMA_RR
    ),
    color = "grey",
    linewidth = 0.4
  ) +
  
  facet_wrap(~ Event,
             scales = "free_x",
             ncol = 1) +
  
  labs(
    title = paste("Visualization: EWMA (Empatica)"),
    subtitle = paste("IBI Time Series:", selected_id),
    x = "Seconds",
    y = "IBI interval (ms)"
  ) +
  
  theme_minimal(base_size = 13) +
  
  theme(
    plot.title = element_text(
      size = 15
    ),
    plot.subtitle = element_text(size=8),
    
    strip.text = element_text(
      size = 10
    ),
    
    panel.grid.minor = element_blank()
  )

```

## 5.2 For Polar

### 5.2.1 Artifact Detection: RR Data 

```{r artefact-detection-for-polar-rr-data}

#Artefact Detection 
Polar_RR_Raw <- Polar_RR_Relevant %>%  
  group_by(ID) %>%
  
  #Absolute Threshold
  mutate(
    Artifact_Flag = case_when(
      RR_Raw < 375 ~ "Artifact",
      RR_Raw > 1200 ~ "Artifact", 
      TRUE ~ "OK" )) %>% 
  
 #Relative Threshold
  mutate(
    Local_Median = slide_dbl(
      .x = RR_Raw,
      .before = 5,  
      .after = 5,   
      .f = ~ median(.x, na.rm = TRUE)),
    RR_Deviation = abs(RR_Raw - Local_Median),
    Artifact_Flag = case_when(
    Artifact_Flag == "Artifact" ~ "Artifact", 
    RR_Deviation > 250 ~ "Artifact", 
    TRUE ~ "OK")) %>%
  
  ungroup()


```

### 5.2.2 Artifact Imputation 

```{r artifcat-imputation-polar-rr}

#Change Flagged Artifacts into NAs
Polar_RR_Flagged <- Polar_RR_Raw %>% 
  mutate(RR_clean = if_else(Artifact_Flag == "Artifact", NA_real_, RR_Raw))


#Impute TS Function: NA_MA replaces NAs w/ moving average
Polar_RR_Imputed <- Polar_RR_Flagged %>%
  group_by(ID) %>%
  arrange(Timestamp) %>%
  group_modify(~ {
    .x$EWMA_RR <- na_ma(.x$RR_clean, k = 32, weighting = "exponential")
    .x
  }) %>%
  ungroup()

```

### 5.2.3 Imputation Visualization

```{r ewma-visualization-polar-rr}

# Select participant
selected_id <- "OITC28"

# Prepare plotting data
plot_data <- Polar_RR_Imputed %>%
  
  filter(ID == selected_id) %>%
  filter (Event %in% c("Trial1", "Drone")) %>% 
  
  #group_by(Event) %>%
  
  arrange(Timestamp) %>%
  
  mutate(
    Seconds = row_number(),
    Imputed = is.na(RR_clean)
  ) %>%
  
  ungroup()

# Plot
ggplot(plot_data,
       aes(x = Seconds,
           y = EWMA_RR)) +
  
  # Main RR signal
  geom_line(color = "blue",
            linewidth = 0.4) +
  
  # Original artifacts
  geom_point(
    data = subset(plot_data, Imputed),
    aes(y = RR_Raw),
    color = "red",
    shape = 4,
    size = 2
  ) +
  
  # EWMA imputed values
  geom_point(
    data = subset(plot_data, Imputed),
    aes(y = EWMA_RR),
    color = "black",
    shape = 17,
    size = 2
  ) +
  
  # Connect raw artifact to imputed value
  geom_segment(
    data = subset(plot_data, Imputed),
    aes(
      xend = Seconds,
      y = RR_Raw,
      yend = EWMA_RR
    ),
    color = "grey",
    linewidth = 0.4
  ) +
  
  facet_wrap(~ Event,
             scales = "free_x",
             ncol = 1) +
  
  labs(
    title = paste("Visualization: EWMA (Polar)"),
    subtitle = paste("RR Time Series:", selected_id),
    x = "Seconds",
    y = "RR interval (ms)"
  ) +
  
  theme_minimal(base_size = 13) +
  
  theme(
    plot.title = element_text(
      size = 15
    ),
    
    plot.subtitle = element_text(size=8),
    
    strip.text = element_text(
      size = 10
    ),
    
    panel.grid.minor = element_blank()
  )

```

# 6. Compute HR

## For Empatica

```{r compute-hr-empatica}

#Compute beat-to-beat HR
Empatica_HR <- Empatica_RR_Imputed %>%
  mutate(
    # Heart Rate from imputed RR
    HR = round(60000 / EWMA_RR)) %>% 
  dplyr:: select(-c(RR_Raw, Artifact_Flag, Local_Median, RR_Deviation, RR_clean, EWMA_RR ))

#Compute to 1-second HR
Empatica_HR <- Empatica_HR %>%
  group_by(ID, Event, timestamp) %>%
  mutate(
    HR = mean(HR, na.rm = TRUE)
  ) %>% 
  ungroup()

```

## For Polar

```{r compute-hr-polar}

#Compute beat-to-beat HR
Polar_HR <- Polar_RR_Imputed %>%
  mutate(
    # Beat-to-beat HR
    HR = 60000 / EWMA_RR,

    # Timestamp with Nanoseconds, convert to Timestamp with Millisecond
    
    ## Convert timestamp to numeric seconds
    Time_numeric = as.numeric(Timestamp),

    ## Millisecond precision
    Time_ms = floor(Time_numeric * 1000) / 1000,

    ## Second bin
    Time_sec = floor(Time_ms)
  )

#Compute to 1-second HR
Polar_HR <- Polar_HR %>%
 group_by(
    ID,
    Session,
    Color,
    Station,
    Event,
    Group,
    Type,
    Time_sec
  ) %>%

  summarise(
    HR = round(mean(HR, na.rm = TRUE)),
    .groups = "drop"
  ) %>%

  mutate(
    Timestamp = hms::as_hms(Time_sec)
  ) %>%

  select(
    ID, Session, Color, Station,
    Timestamp, HR, Event, Group, Type
  )

```

# 7. Normalize HR 

## 7.1 For Empatica 

```{r create-HR-quotient-empatica}

#Create Mean Baseline
Empatica_Base_HR <- Empatica_HR %>%
  filter(Event == "Baseline") %>%
  group_by(ID) %>%
  summarise(Base_HR = round(mean(HR, na.rm = TRUE)), .groups = "drop")

# 2. Join baseline HR to full HR data
Empatica_HR <- Empatica_HR %>% 
  left_join(Empatica_Base_HR, by = "ID")

# 3. Compute Quotient and Difference per row
Empatica_HR <- Empatica_HR %>%
  group_by(ID) %>%  
  mutate(
    HR_Quo = (HR - Base_HR) / Base_HR  ) %>%
  ungroup()


# 4. Create 5-Min Epochs per Event
Empatica_HR_Epochs <- Empatica_HR %>%
  group_by(ID, Session, Event) %>% 
  arrange(timestamp, .by_group = TRUE) %>%
 mutate(
    event_start = min(timestamp),
    elapsed_sec = as.numeric(timestamp - event_start),
    epoch_5min = floor(elapsed_sec / 300) + 1,
    Elapsed_second_epoch = elapsed_sec %% 300,
    Event_Epoch = paste0(Event, "_", epoch_5min)
  ) %>%
  group_by(ID, Event_Epoch) %>%
  filter(
    max(Elapsed_second_epoch) >= 214) %>%
  ungroup()

#5. Compute Mean per Epoch 
Empatica_HR_Epochs <- Empatica_HR_Epochs %>% 
   group_by(ID, Event_Epoch) %>% 
  summarise(
   
     # Keep participant info
    Session = first(Session),
    Color = first(Color),
    Station = first(Station),
    Group = first(Group), 
    Type = first(Type), 
    Event = first(Event),
   
    # Number of observations
    N_HR = n(),
    
     # Quotient
    HR_Quo = round(mean(HR_Quo, na.rm = TRUE), 2),
    
    # HR
    HR = round(mean(HR, na.rm = TRUE)),

    
    .groups = "drop")
  

```

## 7.2 For Polar 

```{r create-HR-quotient-polar}

# 1. Create Mean Baseline
Polar_Base_HR <- Polar_HR %>%
  filter(Event == "Baseline") %>%
  group_by(ID) %>%
  summarise(Base_HR = round(mean(HR, na.rm = TRUE)), .groups = "drop")

# 2. Join baseline HR to full HR data
Polar_HR <- Polar_HR %>% 
  left_join(Polar_Base_HR, by = "ID")

# 3. Compute Quotient and Difference per row
Polar_HR <- Polar_HR %>%
  group_by(ID) %>%  
  mutate(
    HR_Quo = (HR - Base_HR) / Base_HR  ) %>%
  ungroup()


# 4. Create 5-Min Epochs per Event
Polar_HR_Epochs <- Polar_HR %>%
  group_by(ID, Session, Event) %>%
  arrange(Timestamp, .by_group = TRUE) %>%
  mutate(
    event_start = min(Timestamp),
    elapsed_sec = as.numeric(Timestamp - event_start),
    epoch_5min = floor(elapsed_sec / 300) + 1,
    Elapsed_second_epoch = elapsed_sec %% 300,
    Event_Epoch = paste0(Event, "_", epoch_5min)
  ) %>%
  group_by(ID, Event_Epoch) %>%
  filter(
    max(Elapsed_second_epoch) >= 214) %>%
  ungroup()
  

#5. Compute Mean per Epoch 
Polar_HR_Epochs <- Polar_HR_Epochs %>% 
   group_by(ID, Event_Epoch) %>% 
  summarise(
   
     # Keep participant info
    Session = first(Session),
    Color = first(Color),
    Station = first(Station),
    Group = first(Group), 
    Type = first(Type), 
    Event = first(Event),
   
    # Number of observations
    N_HR = n(),
    
     # Quotient
    HR_Quo = round(mean(HR_Quo, na.rm = TRUE), 2),
    
    # HR
    HR = round(mean(HR, na.rm = TRUE)),
    
    .groups = "drop")
  

```

# 7. Time-based Measures

## 7.1 For Empatica 

```{r time-based-measures-empatica}

# Create 5-Min Epochs per Event
Empatica_Epoch <- Empatica_RR_Imputed %>% 
 group_by(ID, Session, Event) %>% 
  arrange(timestamp, .by_group = TRUE) %>%
  mutate(
    event_start = min(timestamp),
    elapsed_sec = as.numeric(timestamp - event_start),
    epoch_5min = floor(elapsed_sec / 301) + 1,
    Elapsed_second_epoch = elapsed_sec %% 301,
    Event_Epoch = paste0(Event, "_", epoch_5min)
  ) %>%
  group_by(ID, Event_Epoch) %>%
  filter(
    max(Elapsed_second_epoch) >= 214) %>%
  ungroup()

 Check <- Empatica_Epoch %>% distinct(ID, Event_Epoch)


#Compute Timebased-metrics per Epoch
Empatica_Timebased_Epoch <- Empatica_Epoch %>% 
     group_by(ID, Event_Epoch) %>% 
  summarise(
    
    # Keep participant info
    Session = first(Session),
    Color = first(Color),
    Station = first(Station),
    Group = first(Group), 
    Type = first(Type), 
    Event = first(Event), 
    
    #Count RR
    N_RR = n(),
    
    # SDNN
    SDNN = round(sd(EWMA_RR, na.rm = TRUE), 2),
    
    # log SDNN
    lnSDNN = round(log(
      sd(EWMA_RR, na.rm = TRUE)
    ), 3),
    
    # RMSSD
    RMSSD = round(
      sqrt(
        mean(
          diff(EWMA_RR)^2,
          na.rm = TRUE
        )
      ),
      2
    ),
    
    # log RMSSD
    lnRMSSD = round(log(
      sqrt(
        mean(
          diff(EWMA_RR)^2,
          na.rm = TRUE
        )
      )
    ), 3),

    
    .groups = "drop"
  ) 

```

## 7.2 For Polar 

```{r time-based-measures-empatica}

#Create 5minutes epoch per event 
Polar_Epoch <- Polar_RR_Imputed %>%
  group_by(ID, Session, Event) %>%
  arrange(Timestamp, .by_group = TRUE) %>%
  mutate(
    event_start = min(Timestamp),
    elapsed_sec = as.numeric(Timestamp - event_start),
    epoch_5min = floor(elapsed_sec / 301) + 1,
    Elapsed_second_epoch = elapsed_sec %% 301,
    Event_Epoch = paste0(Event, "_", epoch_5min)
  ) %>%
  group_by(ID, Event_Epoch) %>%
  filter(max(Elapsed_second_epoch) >= 214)%>%
  ungroup()

 Check <- Polar_Epoch %>% distinct(ID, Event_Epoch)

#Compute Timebased-metrics per Epoch
Polar_Timebased_Epoch <- Polar_Epoch %>% 
     group_by(ID, Event_Epoch) %>% 
  summarise(
    
    # Keep participant info
    Session = first(Session),
    Color = first(Color),
    Station = first(Station),
    Group = first(Group), 
    Type = first(Type), 
    Event = first(Event),
    
    #Count RR
    N_RR = n(),
    
    # SDNN
    SDNN = round(sd(EWMA_RR, na.rm = TRUE), 2),
    
    # log SDNN
    lnSDNN = round(log(
      sd(EWMA_RR, na.rm = TRUE)
    ), 3),
    
    # RMSSD
    RMSSD = round(
      sqrt(
        mean(
          diff(EWMA_RR)^2,
          na.rm = TRUE
        )
      ),
      2
    ),
    
    # log RMSSD
    lnRMSSD = round(log(
      sqrt(
        mean(
          diff(EWMA_RR)^2,
          na.rm = TRUE
        )
      )
    ), 3),

    
    .groups = "drop"
  ) 

# Test: lnRMSSD_Quo

# 1. Mean baseline lnRMSSD per participant
Polar_Base_lnRMSSD <- Polar_Timebased_Epoch %>%
  filter(Event == "Baseline") %>%
  group_by(ID) %>%
  summarise(
    Base_lnRMSSD = mean(lnRMSSD, na.rm = TRUE),
    Base_RMSSD = mean(RMSSD, na.rm = TRUE), 
    .groups = "drop"
  )

# 2. Join baseline to all epochs
Polar_Timebased_Epoch <- Polar_Timebased_Epoch %>%
  left_join(Polar_Base_lnRMSSD, by = "ID")

# 3. Compute quotient
Polar_Timebased_Epoch <- Polar_Timebased_Epoch %>%
  mutate(
    lnRMSSD_Quo = (lnRMSSD - Base_lnRMSSD) / Base_lnRMSSD, 
    RMSSD_Quo = (RMSSD - Base_RMSSD) / Base_RMSSD
  )


```

# 8. Frequency-based Measures

Done in Python

Export RR Data 

```{r export-RR-data}
#Set WD
setwd("./Preprocessed_Data")

#Export Imputed Timeseries 
write_xlsx(Empatica_Epoch, "Empatica_Epoch.xlsx")
write_xlsx(Polar_Epoch, "Polar_Epoch.xlsx")

```


Import from Python :)

```{r import-hrv-data-from-python}

setwd("./Preprocessed_Data")

#Empatica
Empatica_Frequency_HRV <- read_excel("Empatica_Frequency_HRV.xlsx")
view(Empatica_Frequency_HRV)
#Polar
Polar_Frequency_HRV <- read_excel("Polar_Frequency_HRV.xlsx")

```

# 9.  Unify Aggregated Data 

## HRV Metrics 

```{r unify-aggregated-data-rr}

#Empatica
HRV_Summary_Empatica <- Empatica_Timebased_Epoch %>% 
  left_join(Empatica_Frequency_HRV, by =c("ID", "Event_Epoch")) %>% 
  rename_with(
    ~ paste0(., "_Empatica"),
    .cols = c(SDNN, lnSDNN, RMSSD, lnRMSSD, N_RR, HRV_LF, HRV_HF, HRV_LFHF, HRV_HFn, HRV_LFn ))  %>% 
  dplyr::select(-c(14,15,18,19,23))


#Polar
HRV_Summary_Polar <- Polar_Timebased_Epoch %>%
  left_join(Polar_Frequency_HRV, by =c("ID", "Event_Epoch")) %>% 
  rename_with(
    ~ paste0(., "_Polar"),
    .cols = c(SDNN, lnSDNN, RMSSD, lnRMSSD, lnRMSSD_Quo, RMSSD_Quo , N_RR, HRV_LF, HRV_HF, HRV_LFHF,HRV_HFn, HRV_LFn))   %>% 
  dplyr::select(-c(14,15,18,19,23))


#Join both Devices 
HRV_Comparison <- HRV_Summary_Empatica %>% 
  left_join(HRV_Summary_Polar, by=c("ID", "Event", "Event_Epoch", "Group", "Session", "Color", "Station", "Type"))

```

## HR Metrics 

```{r unify-aggregated-data-hr}

#Empatica
HR_Summary_Empatica <- Empatica_HR_Epochs %>%
  rename_with(
    ~ paste0(., "_Empatica"),
    .cols = c(HR_Quo,HR, N_HR)
  )


#Polar
HR_Summary_Polar <- Polar_HR_Epochs %>%
  rename_with(
    ~ paste0(., "_Polar"),
    .cols = c(HR_Quo,HR, N_HR)
  )



#Join both Devices 
HR_Comparison <- HR_Summary_Empatica %>% 
  left_join(HR_Summary_Polar, by=c("ID", "Event", "Event_Epoch", "Group", "Session", "Color", "Station", "Type"))



```

## Create Cardio Summary 

```{r cardio-summary-}

#Join 
Cardio_Comparison <- data.frame()
Cardio_Comparison <- HRV_Comparison %>% 
  left_join(HR_Comparison, by=c("ID", "Event", "Event_Epoch", "Group", "Session", "Color",
                                "Station", "Type"))

```


# 10. Export Data 

## 10.1 Create a new grouping variable

```{r}

#Cardio_Comparison
Cardio_Comparison <- Cardio_Comparison %>%
  mutate(
    Event_Phase = case_when(
      Event %in% c("Trial1", "Trial2", "Trial3", "Video") ~ "Trials",
      TRUE ~ Event) ) 


#Empatica RR Imputed
Empatica_RR_Imputed <- Empatica_RR_Imputed %>%
  mutate(
    Event_Phase = case_when(
      Event %in% c("Trial1", "Trial2", "Trial3", "Video") ~ "Trials",
      TRUE ~ Event) ) 

#Empatica_HR
Empatica_HR <- Empatica_HR %>%
  mutate(
    Event_Phase = case_when(
      Event %in% c("Trial1", "Trial2", "Trial3", "Video") ~ "Trials",
      TRUE ~ Event) ) 

#Polar RR Imputed
Polar_RR_Imputed <- Polar_RR_Imputed %>%
  mutate(
    Event_Phase = case_when(
      Event %in% c("Trial1", "Trial2", "Trial3", "Video") ~ "Trials",
      TRUE ~ Event) ) 

#Polar_HR
Polar_HR <- Polar_HR %>%
  mutate(
    Event_Phase = case_when(
      Event %in% c("Trial1", "Trial2", "Trial3", "Video") ~ "Trials",
      TRUE ~ Event) ) 

```

## 10.2 Aggregated Values and Imputed Timeseries

```{r export-aggregated values}

# Set Directory 
setwd("./Preprocessed_Data")


## Aggregated Dataset 
write_xlsx(Cardio_Comparison, "Cardio_Comparison.xlsx")

## Timeseries
write_xlsx(Empatica_RR_Imputed, "Empatica_RR_Imputed.xlsx")
write_xlsx(Empatica_HR, "Empatica_HR.xlsx")
write_xlsx(Polar_RR_Imputed, "Polar_RR_Imputed.xlsx")
write_xlsx(Polar_HR, "Polar_HR.xlsx")

```














